---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

    The source of this data is the Million Song Dataset.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

There were several curators for this dataset---the Million Song Dataset used a company called the Echo Nest to derive data points about one million contemporary songs. It was also a collaboration between the Echo Nest and LabROSA (laboratory working towards intelligent machine listening. It was collected in 2011.\*\*

-   Write a brief description of the observations.

A brief description of the observations in this data set is that there are different songs associated by a song ID and their respective artists that also have an artist ID.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

How does the relationship between tempo and song popularity differ by the genre associated with the artist (artist.terms)?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

We will take a look at the relationship between the tempos of songs and their popularity by four specific genres (rap, techno, country, rock) associated with the artists. After looking into this relationship, we will be able to establish if certain genres are more or less popular due to faster or slower tempos. We predict that rock and techno songs will have a strong positive relationship between tempo and popularity and that rap and country will still have a positive relationship, but not a strong one.

-   Identify the types of variables in your research question. Categorical? Quantitative?

The categorical variable is genre associated with artist (artist.terms). The quantitative variables are tempo and popularity.

Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

Our research question builds on this article because we are similarly looking into audio features (like tempo) that make songs more popular but going further to see if the genre (not an audio feature) reveals a relationship.

## Glimpse of data

```{r}
#| label: load-data-1

music <- read_csv("data/music.csv")

glimpse(music)
```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

# add code here
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

# add code here
```
