---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

    The source of this data is the Million Song Dataset.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    There were several curators for this dataset---the Million Song Dataset used a company called the Echo Nest to derive data points about one million contemporary songs. It was also a collaboration between the Echo Nest and LabROSA (laboratory working towards intelligent machine listening. It was collected in 2011.\*\*

-   Write a brief description of the observations.

    A brief description of the observations in this data set is that there are different songs associated by a song ID and their respective artists that also have an artist ID.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    How does the relationship between tempo and song popularity differ by the genre associated with the artist (artist.terms)?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    We will take a look at the relationship between the tempos of songs and their popularity by four specific genres (rap, techno, country, rock) associated with the artists. After looking into this relationship, we will be able to establish if certain genres are more or less popular due to faster or slower tempos. We predict that rock and techno songs will have a strong positive relationship between tempo and popularity and that rap and country will still have a positive relationship, but not a strong one.

-   Identify the types of variables in your research question. Categorical? Quantitative?The categorical variable is genre associated with artist (artist.terms). The quantitative variables are tempo and popularity.

## Literature

-   Find one published credible article on the topic you are interested in researching.

    Nijkamp, R. (2018). Prediction of product success: explaining song popularity by audio features from Spotify data (Bachelor's thesis, University of Twente). http://essay.utwente.nl/75422/1/NIJKAMP_BA_IBA.pdf

-   Provide a one paragraph summary about the article.

    *Prediction of product success: explaining song popularity by audio features from Spotify data* attempts to explore the relation between song attributes such as tempo and key signature, as well as the number of plays on Spotify, the popular music-streaming platform. This paper uses a dataset of 1000 songs to investigate this question. Ultimately, the researchers find that there is only a fairly limited relation between song popularity (measured in stream count) and song features. In particular, the paper finds that tempo, which is the variable that we are most interested in, has no significant relation to stream count.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

    Our research question builds on this article because we are similarly looking into audio features (like tempo) that make songs more popular but going further to see if a relation can be found conditional upon the genre (not an audio feature) associated with the artist a song.

## Glimpse of data

```{r}
#| label: load-data-1

music <- read_csv("data/music.csv")

glimpse(music)
```

# Data 2

## Introduction and data

-   Identify the source of the data.

    This data came from the Coffee Quality Institute's review pages. 

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The data was collected in January 2018 by James LeDoux who created the Coffee Quality Database and pulled variables from review pages. 

-   Write a brief description of the observations.

    The Coffee Ratings dataset has 990 observations with each observation being a review of coffee beans by one of the Coffee Quality Institute's reviewers. The reviewer rates the coffee based upon different quality measures like flavor, aroma, and sweetness. Each observation also has information about the coffee bean like color and farm data like where it was grown, altitude grown at and region. 

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    Does the species or the average altitude of where coffee was grown have a greater impact on the flavor score given to coffee by a reviewer?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    Within this topic, we are researching whether just the species influences a better flavor score on a scale from 0-10 or if average altitude of where the coffee was grown influences a better flavor score on a scale from 0-10. This will help us determine if a certain species leads reviewers to rank the flavor score higher or if the altitude of where a species was grown changes the flavor score to be higher or lower. Our hypothesis is that species alone, no matter the altitude, influences a greater flavor score.

-   Identify the types of variables in your research question. Categorical? Quantitative?

    Species is a categorical variable. Average altitude and flavor score are quantitative variables.

## Literature

-   Find one published credible article on the topic you are interested in researching.

    Ana Paula Pereira Bressani, Silvia Juliana Martinez, Nádia Nara Batista, João Batista Pavesi Simão, Disney Ribeiro Dias, Rosane Freitas Schwan, Co-inoculation of yeasts starters: A strategy to improve quality of low altitude Arabica coffee, Food Chemistry, Volume 361, 2021, 130133, ISSN 0308-8146,<https://doi.org/10.1016/j.foodchem.2021.130133>. 

-   Provide a one paragraph summary about the article.

    This article is examining how the quality of coffee of the species Arabica could be improved when grown at low altitudes. The scope of the article discusses using yeast inoculation but also describes the impact of altitude on coffee flavor in a specific species. It states that Arabica is a very popular species due to its sweet flavor elements. Flavor is determined by volatile compounds in the coffee which is influenced by factors like geographical origin, climate, altitude, species, harvesting method, processing and storage. The researchers have determined altitude to be an extremely important piece of determining flavor, especially at lower altitude. 

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

    Our research question builds on the article as we are examining more species than just Arabica and determining if altitude is the most important factor instead of just an important factor by looking at the average of altitude and not just low altitude.

## Glimpse of data

```{r}
#| label: load-data-2

coffee <- read_csv("data/coffee.csv")


glimpse(coffee)
```

# Data 3

## Introduction and data

-   Identify the source of the data.

    The source of the data is the Video Games CSV File from the CORGIS Datasets.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    It was originally collected in 2011 by Dr. Joe Cox from the crowd-sourced data on "How Long to Beat". The date for the lifetime unit sales of games released variable was collected from an online database maintained by VGChartz. 

-   Write a brief description of the observations.

    There are over one thousand video games reported in this data set, and measure a variety of things from sales performance to game playing time to console type. Some of these variables, like console type, are categorical while average time spent playing the main game storyline is quantitative.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    Are factors like average time spent playing the entire game or whether a game is available on multiple platforms associated with the sales performance of a video game, and if so, what variable has the largest correlation?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    Our research topic considers how different factors, like time spent playing a video game or its availability, affect the game's sales performance. Total sales of the game is measured in millions of dollars, availability is measured on whether the game is available on multiple platforms, with a value of "True" if it is and "False" if it isn't, and time is measured in hours. Our hypothesis is ??????????? NOT FNISHED HERE NOT FINISHED HERE

-   Identify the types of variables in your research question. Categorical? Quantitative?

    Average time (and all variables measuring time): Quantitative, continuous 

    Sales performance: Quantitative, continuous 

    If a game is available on multiple platforms: Categorical, true or false

## Literature

-   Find one published credible article on the topic you are interested in researching.

    "Uncertainty in the Movie Industry: Does Star Power Reduce the Terror of the Box Office?" by De Vany, Arthur and Walls, W. David

-   Provide a one paragraph summary about the article.

    "Uncertainty in the Movie Industry: Does Star Power Reduce the Terror of the Box Office?" by De Vany, Arthur and Walls, W. David, explores the role of movie stars in reducing uncertainty in the movie industry. The authors examine the relationship between the presence of "stars" or A-List actors in movies, and their box office performance. Their findings are that star power can reduce the uncertainty associated with box office success, but only up to a certain point - their ability to influence a film's performance in the box office is limited. The study also highlights the importance of other factors such as genre, quality, and marketing in determining box office success. Overall, the paper provides insights into the complex nature of the movie industry and the factors that influence moviegoers' decision-making.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

    While the article examines how factors of a movie can impact its sales success in the box office, our data and research question are concerned with video games. Thus, we can apply some of the same analytical methods to examining our data set, while keeping in mind all the ways video games are different, including the way consumers interact with the product, the demographic of consumers, and more. 

## Glimpse of data

```{r}
#| label: load-data-3

videogame <- read_csv("data/video_games.csv")
glimpse(videogame)

```
